Sequential rank agreement methods for comparison of ranked lists

### 2015-05-05 latex/paper.Rnw
Figure 1

\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[]{graphicx}
\usepackage{amsmath}
\usepackage[]{color}
\usepackage{subcaption}
\usepackage{natbib}
\usepackage{xspace}
\usepackage[noae]{Sweave}

@@

\title{Sequential rank agreement methods for comparison of ranked lists}
\author{Claus Thorn Ekstrøm and Thomas Alexander Gerds and Kasper Brink-Jensen}

@@

\subsection{Agreement among fully observed lists}
We shall start by the simplest case where all $L$ lists are fully
observed, \ie, we have the rank of all $P$ items for all of the $L$
lists. Fully observed ranked lists are common and occur, for example,
when we have a single dataset and apply different statistical analysis
methods to produce lists of predictors ranked according to their
importance, or if the same analysis method is applied to data from
different populations.

For the fully obseved list situation we can plot the sequential rank
agreement \eqref{def:sra} as a function of depth $d$.  An example is
seen in Figure~\ref{fig:example1} where smaller values on the $y$ axis
correspond to better agreement among the lists than larger values. The
sequential rank agreement at depth $d$ can be interpreted as how many
ranks apart do we on average expect the items found until depth $d$ to
be.

If the lists generally agree on all items then we expect the
sequential rank agreement curve to be flat, but if there are changes
in the rank agreement levels then this suggests that there are changes
in the levels of agreement. In particular, we expect the agreement
among the lists to be better towards the top of the lists and worse
towards the end of the lists.

\begin{figure}[tb]
\begin{center}
\includegraphics[width=.45\textwidth]{paper-fig1a}
\includegraphics[width=.45\textwidth]{paper-fig1b}
\end{center}
 \caption{Sequential rank agreement for 4 different analysis methods
   applied to the Golub data (thick red line). The red and blue shaded
 areas represent the pointwise 95\% normal areas for the XXX and XXY benchmarks,
respectively.}
 \label{fig:example1}
\end{figure}

@@

\begin{figure}[tb]
\begin{center}
\includegraphics[width=.45\textwidth]{paper-fig2a}
\includegraphics[width=.45\textwidth]{paper-fig2b}
\end{center}
 \caption{Sequential rank agreement for 4 different analysis methods
   applied to the Golub data (thick red line). The red and blue shaded
 areas represent the pointwise 95\% normal areas for the XXX and XXY benchmarks,
respectively.}
 \label{fig:example1}
\end{figure}

>>>>

\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[]{graphicx}
\usepackage{amsmath}
\usepackage[]{color}
\usepackage{subcaption}
\usepackage{natbib}
\usepackage{xspace}
\usepackage[noae]{Sweave}
\usepackage{authblk}

@@

\title{Sequential rank agreement methods for comparison of ranked
  lists}
\author{Claus Thorn Ekstrøm}
\author{Thomas Alexander Gerds}
\author{Kasper Brink-Jensen}
\affil{Biostatistics, University of Copenhagen}

% Use square brackets for numbering affiliations

@@

\subsection{Agreement among fully observed lists}
The simplest case occurs when all $L$ lists are fully
observed, \ie, we have observed the rank of all $P$ items for all $L$
lists. Fully observed ranked lists are common and arise, for example,
when different statistical analysis methods are applied to a single
dataset to produce lists of predictors ranked according to their
importance, or if the same analysis method is applied to data from
different populations.

With fully observed lists we can plot the sequential rank agreement
\eqref{def:sra} as a function of depth $d$.  An example is seen in the
left panel of Figure~\ref{fig:example1} where four analysis methods
were used to rank genes (for further detail see the application
section below). The sequential rank agreement at depth $d$ can be
interpreted as how many ranks apart do we on average expect the items
found until depth $d$ to be so smaller values on the $y$ axis
correspond to better agreement among the lists than larger values.

If the lists generally agree on all items then we expect the
sequential rank agreement curve to be flat, but if there are changes
in the levels of rank agreement then this suggests that there are changes
in the levels of agreement. In particular, we generally expect the
agreement among the lists to be better towards the top of the lists
and worse towards the end of the lists, and in this case the sequential rank
agreement curve will start at a low level, and then have an increase
before it levels off.

\begin{figure}[tb]
\begin{center}
\includegraphics[width=.45\textwidth]{paper-fig1a}
\includegraphics[width=.45\textwidth]{paper-fig1b}
\end{center}
 \caption{Left panel: Sequential rank agreement for 4 different analysis methods
   applied to the 3051 genes in the Golub data (solid line). The dashed line shows the
   corresponding estimated sequential rank agreement for censored
   lists based purely on the top 50 ranked genes. Right panel contains
   the same information but has zoomed in on the first part of the $x$-axis.}
 \label{fig:example1}
\end{figure}

@@

\begin{figure}[tb]
\begin{center}
\includegraphics[width=.45\textwidth]{paper-fig2a}
\includegraphics[width=.45\textwidth]{paper-fig2b}
\end{center}
 \caption{Sequential rank agreement for 4 different analysis methods
   applied to the Golub data (thick red line). The red and blue shaded
 areas represent the pointwise 95\% normal areas for the XXX and XXY benchmarks,
respectively.}
 \label{fig:case1}
\end{figure}
